% Copyright (C) 2014-2016 by Thomas Auzinger <thomas@auzinger.name>

\documentclass[draft,final]{vutinfth} % Remove option 'final' to obtain debug information.

% Load packages to allow in- and output of non-ASCII characters.
\usepackage{lmodern}        % Use an extension of the original Computer Modern font to minimize the use of bitmapped letters.
\usepackage[T1]{fontenc}    % Determines font encoding of the output. Font packages have to be included before this line.
\usepackage[utf8]{inputenc} % Determines encoding of the input. All input files have to use UTF8 encoding.

% Extended LaTeX functionality is enables by including packages with \usepackage{...}.
\usepackage{amsmath}    % Extended typesetting of mathematical expression.
\usepackage{amssymb}    % Provides a multitude of mathematical symbols.
\usepackage{mathtools}  % Further extensions of mathematical typesetting.
\usepackage{microtype}  % Small-scale typographic enhancements.
\usepackage[inline]{enumitem} % User control over the layout of lists (itemize, enumerate, description).
\usepackage{multirow}   % Allows table elements to span several rows.
\usepackage{booktabs}   % Improves the typesettings of tables.
\usepackage{subcaption} % Allows the use of subfigures and enables their referencing.
\usepackage[ruled,linesnumbered,algochapter]{algorithm2e} % Enables the writing of pseudo code.
\usepackage[usenames,dvipsnames,table]{xcolor} % Allows the definition and use of colors. This package has to be included before tikz.
\usepackage{nag}       % Issues warnings when best practices in writing LaTeX documents are violated.
\usepackage{todonotes} % Provides tooltip-like todo notes.
\usepackage{rotating}  % Support for sideways figures
\usepackage{hyperref}  % Enables cross linking in the electronic document version. This package has to be included second to last.
\usepackage[acronym,toc]{glossaries} % Enables the generation of glossaries and lists fo acronyms. This package has to be included last.

% Define convenience functions to use the author name and the thesis title in the PDF document properties.
\newcommand{\authorname}{Robert Thurnher} % The author name without titles.
\newcommand{\thesistitle}{TempMunger} % The title of the thesis. The English version should be used, if it exists.

\DeclareMathOperator*{\argmin}{arg\,min}

% Set PDF document properties
\hypersetup{
    pdfpagelayout   = TwoPageRight,           % How the document is shown in PDF viewers (optional).
    linkbordercolor = {Melon},                % The color of the borders of boxes around crosslinks (optional).
    pdfauthor       = {\authorname},          % The author's name in the document properties (optional).
    pdftitle        = {\thesistitle},         % The document's title in the document properties (optional).
    pdfsubject      = {TempMunger},              % The document's subject in the document properties (optional).
    pdfkeywords     = {data wrangling, visual analytics, visual data transformation} % The document's keywords in the document properties (optional).
}

\setpnumwidth{2.5em}        % Avoid overfull hboxes in the table of contents (see memoir manual).
\setsecnumdepth{subsection} % Enumerate subsections.

\nonzeroparskip             % Create space between paragraphs (optional).
\setlength{\parindent}{0pt} % Remove paragraph identation (optional).

\makeindex      % Use an optional index.
\makeglossaries % Use an optional glossary.
%\glstocfalse   % Remove the glossaries from the table of contents.

% Set persons with 4 arguments:
%  {title before name}{name}{title after name}{gender}
%  where both titles are optional (i.e. can be given as empty brackets {}).
\setauthor{}{\authorname}{Bakk.}{male}
\setadvisor{Univ.Prof. Dr.}{Silvia Miksch}{}{female}

% For bachelor and master theses:
\setfirstassistant{Dr.}{Theresia Gschwandtner}{}{female}
%\setsecondassistant{Pretitle}{Forename Surname}{Posttitle}{male}
%\setthirdassistant{Pretitle}{Forename Surname}{Posttitle}{male}

% For dissertations:
%\setfirstreviewer{Pretitle}{Forename Surname}{Posttitle}{male}
%\setsecondreviewer{Pretitle}{Forename Surname}{Posttitle}{male}

% For dissertations at the PhD School and optionally for dissertations:
%\setsecondadvisor{Pretitle}{Forename Surname}{Posttitle}{male} % Comment to remove.

% Required data.
\setaddress{Neubaugasse 64-66, A-1070 Wien; \textcolor{blue}{\href{mailto:robert@thurnher.email}{robert[at]thurnher.email}}}
\setregnumber{0004297}
\setdate{13}{04}{2017} % Set date with 3 arguments: {day}{month}{year}.
\settitle{\thesistitle}{TempMunger} % Sets English and German version of the title (both can be English or German).
\setsubtitle{A Visual Analytics Approach Supporting Transformations of Time-Oriented Data}{Ein Visual Analytics Ansatz zur Transformation Zeitorientierter Daten} % Sets English and German version of the subtitle (both can be English or German).

% Select the thesis type: bachelor / master / doctor / phd-school.
% Bachelor:
%\setthesis{bachelor}
%
% Master:
\setthesis{master}
\setmasterdegree{dipl.} % dipl. / rer.nat. / rer.soc.oec. / master
%
% Doctor:
%\setthesis{doctor}
%\setdoctordegree{rer.soc.oec.}% rer.nat. / techn. / rer.soc.oec.
%
% Doctor at the PhD School
%\setthesis{phd-school} % Deactivate non-English title pages (see below)

% For bachelor and master:
\setcurriculum{Information \& Knowledge Management}{Information \& Knowledge Management} % Sets the English and German name of the curriculum.

% For dissertations at the PhD School:
%\setfirstreviewerdata{Affiliation, Country}
%\setsecondreviewerdata{Affiliation, Country}


\begin{document}

\frontmatter % Switches to roman numbering.
% The structure of the thesis has to conform to
%  http://www.informatik.tuwien.ac.at/dekanat

\addtitlepage{naustrian} % German title page (not for dissertations at the PhD School).
\addtitlepage{english} % English title page.
\addstatementpage

%\begin{danksagung*}
%\todo{Ihr Text hier.}
%\end{danksagung*}

\newcommand{\heart}{\ensuremath\heartsuit}

\begin{acknowledgements*}
Dedicated to my \textbf{father}\footnote{\emph{\textbf{RIP.}}}, my \textbf{mother}, and my \textbf{wife}. \\
This wouldn't have been possible if it weren't for you. \\
May it also motivate my \textbf{sister} pursuing her dreams. \\
\heart \\

I'd like to thank \emph{Theresia Gschwandtner} for her supervision, understanding, and patience.

Further thanks go out to my \textbf{brother} and \emph{Michael Jakl} for their support and inspiration.

Generally, I'm grateful to everyone who has test-driven the prototype, proofread drafts, and/or provided feedback -- particularly, my \textbf{sister-in-law} and \emph{CVAST} research staff.

\textbf{Thank you!}
\end{acknowledgements*}

\begin{kurzfassung}
\emph{Data-Wrangling} ist im Allgemeinen die mühevolle Aufbereitung von Daten um diese für nachfolgende Analysen nutzbar zu machen.
Normalerweise bedeutet das handgefertigte Skripte anzuwenden, was einen gewissen Grad an technischer Expertise verlangt.
Um derartiges Aufbereiten von Daten auch für User, die eher einer ``Casual''-Kategorie zuordenbar sind, zugänglich zu machen haben wir zur Erleichterung damit verbundener Aufgaben, im Kontext dieser Arbeit einen \emph{Visual-Analytics}-Prototypen erstellt.

Unser Ansatz ist ein interdisziplinärer, der zeitgemäße Konzepte und Ideen aus diversen Gebieten kombiniert:
\emph{Mensch-Computer-Interaktion} und \emph{Usability-Engineering} mit \emph{Information-Retrieval}, \emph{Data-Mining}, \emph{Machine-Learning} sowie \emph{Informationsvisualisierung}.

In diesem Zusammenhang konzentrieren wir uns speziell auf oftmals nötige Transformationsschritte von zeitorientierten Daten.
Diese Art von Daten weist einzigartige Charakteristika auf, die bei verschiedenen Lösungsansätzen gesondert berücksichtigt werden müssen.
Basierend auf den Ergebnissen einer Analyse des State-of-the-Art in diesem Bereich, haben wir offene Fragestellungen sowie einige Anforderungen für einen solchen Prototypen abgeleitet und folglich einen Software-Prototyp namens \textbf{\emph{TempMunger}} in einer agilen Art und Weise iterativ sowohl designt als auch entwickelt.
Der Prozess sowie dazugehörige Artefakte werden umfassend dokumentiert und dargestellt.

Unser Prototyp ist eine web-basierte Anwendung, die für den Gebrauch mittels Desktop-Browser zugeschnitten ist.
Genauer gesagt, bietet sie interaktive Dashboard-Visualisierung-en für möglichst intuitive sowie explorative Transformationsoperationen zeitorientierter Daten.
Eine qualitative Evaluierung des Prototypen zeigte, dass diese Funktionalität als nützlich empfunden wird.
Außerdem ergaben sich verschiedene Ideen für künftige, weiterführende Arbeiten.

Abschließend kann man sagen, dass Data-Wrangling ein spannendes Forschungsgebiet bleibt, in dem es noch viel zu entdecken gilt.
\end{kurzfassung}

\begin{abstract}
\emph{Data wrangling} generally denotes the cumbersome task of making data useful for analysis.
Usually, this means applying hand-crafted scripts, requiring at least a certain degree of technical expertise.
In order to make suchlike data preparation accessible for rather casual users as well, we have built a \emph{visual analytics} prototype in the context of this thesis, easing related tasks.

Our approach is an interdisciplinary one, combining contemporary concepts and ideas from various fields:
\emph{human-computer interaction} and \emph{usability engineering} with \emph{information retrieval}, \emph{data mining}, \emph{machine learning}, plus \emph{information visualization}.

In particular we focus on supporting transformations of time-oriented data since this kind of data exhibits unique characteristics which demand for special consideration.
After analyzing related state of the art we identified open issues and derived requirements for such a system. We followed an iterative design process to develop a software prototype called \textbf{\emph{TempMunger}}, in an agile manner.
The process as well as corresponding artifacts are documented and presented in this thesis.

Our prototype is a web-based application, tailored for desktop browser usage.
More concretely, it offers interactive dashboard visualizations for preferably intuitive and exploratory transformation operations of time-oriented data.
A qualitative evaluation of the prototype demonstrates its usefulness and reveals opportunities for future work.

Concluding it is safe to say that data wrangling continues to be an exciting field of research where much is yet to be discovered.
\end{abstract}

% Select the language of the thesis, e.g., english or naustrian.
\selectlanguage{english}

% Add a table of contents (toc).
\tableofcontents % Starred version, i.e., \tableofcontents*, removes the self-entry.

% Switch to arabic numbering and start the enumeration of chapters in the table of content.
\mainmatter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Introduction}
\label{ch:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/introduction}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{State of the Art}
\label{ch:state-of-the-art}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/state-of-the-art}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Design and Implementation}
\label{ch:design+impl}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/design+impl}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Critical Reflection}
\label{ch:reflection}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/reflection}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Summary and Future Work}
\label{ch:summary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/summary}


\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Software Design and Architecture}
\label{ch:appendix-a}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/appendix-a}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Information Retrieval Background}
\label{ch:appendix-b}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/appendix-b}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Supported Date/Time Formats}
\label{ch:appendix-c}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/appendix-c}


\backmatter

% Use an optional list of figures.
\listoffigures % Starred version, i.e., \listoffigures*, removes the toc entry.

% Use an optional list of tables.
\cleardoublepage % Start list of tables on the next empty right hand page.
\listoftables % Starred version, i.e., \listoftables*, removes the toc entry.

% Use an optional list of alogrithms.
\listofalgorithms
\addcontentsline{toc}{chapter}{List of Algorithms}

% Add an index.
\printindex

% Add a glossary.
\printglossaries

% Add a bibliography.
\bibliographystyle{alpha}
\bibliography{references}

\end{document}
